---
title: "Codebook for Getting and Cleaning Data Course Project"
author: "John Kenneth Velonta"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The first step was to load the features from the features.txt file, which contains the names of the features/variables that we will be utilizing in the assignment.

```{r}
library(tidyverse)

features <- read.delim('./UCI HAR Dataset/features.txt', header=FALSE)
```

The next step was to convert the previously loaded features to a vector for further use.

```{r}
features_char <- unlist(features[['V1']], use.names = FALSE)
```

Next, I loaded the subject data for the test and training datasets. Then I used rbind to merge the test and train subjects.

```{r}
subject_test <- read.delim('./UCI HAR Dataset/test/subject_test.txt', header=FALSE, col.names = "Subject")
subject_train <- read.delim('./UCI HAR Dataset/train/subject_train.txt', header=FALSE, col.names = "Subject")
merged_subjects <- rbind(subject_test, subject_train)
```

Next, I loaded the activities file, merged them through rbind, and changed from numeric to verbose activities through a left_join.

```{r}
y_test <- read.delim('./UCI HAR Dataset/test/y_test.txt', header=FALSE, col.names = "Activity")
y_train <- read.delim('./UCI HAR Dataset/train/y_train.txt', header=FALSE, col.names = "Activity")
Merged_activities <- rbind(y_test,y_train)
activity_labels <- read.delim('./UCI HAR Dataset/activity_labels.txt', header=FALSE)
activity_labels <- cbind(activity_labels, do.call(rbind, strsplit(activity_labels$V1, "( +)")))
names(activity_labels) = c('V1', 'Activity', 'V2')
activity_labels <- mutate(activity_labels, Activity= as.numeric(activity_labels$Activity))
Merged_activities <- left_join(Merged_activities, activity_labels, by=('Activity'))
Merged_activities <- Merged_activities[3]
names(Merged_activities) <- 'Activity'

```

Next, the test and training datasets were loaded and merged through rbind.

```{r}
X_test <- read.delim('./UCI HAR Dataset/test/X_test.txt', header=FALSE)
X_train <- read.delim('./UCI HAR Dataset/train/X_train.txt', header=FALSE)
Merged_data <- rbind(X_test,X_train)
```
I used a combination of strsplit, rbind, and cbind, to separate each feature from each observation to different columns.

```{r}
Merged_data_split <- cbind(Merged_data, do.call(rbind, strsplit(Merged_data$V1, "( +)")))

```
Then I had to deselect columns 1 and 2.

```{r}
Merged_data_split <- Merged_data_split[c(3:563)]

```

Then selected the columns representing the means and standard deviations with the use of grep. Any column name with mean or std on them were selected.

```{r}
Merged_data_split[] <- sapply(Merged_data_split, as.numeric)
names(Merged_data_split) <- features_char
mean_stdev_loc <- grep("mean|std",names(Merged_data_split))
Merged_data_mean_stdev <- Merged_data_split[mean_stdev_loc]

```
Then, for tidying the feature names, f were changed to frequency, t were changed to time, and the numbers parentheses were removed through sub.

```{r}
names(Merged_data_mean_stdev) <- sub("[0-9]*[ ]*f","frequency",names(Merged_data_mean_stdev))
names(Merged_data_mean_stdev) <- sub("[0-9]*[ ]*t","time",names(Merged_data_mean_stdev))
names(Merged_data_mean_stdev) <- sub("[0-9]*[ ]*","",names(Merged_data_mean_stdev))
names(Merged_data_mean_stdev) <- sub("\\(\\)","",names(Merged_data_mean_stdev))
names(Merged_data_mean_stdev) <- tolower(names(Merged_data_mean_stdev))


```

I then merged the datasets, activities, and subjects through cbind.

```{r}
merged_all <- cbind(merged_subjects, Merged_activities, Merged_data_mean_stdev)

```

The resulting merged dataframe was then grouped by subject and activity, and then summarized to get the mean for each grouping.

```{r}
merged_all <- merged_all %>% group_by(Subject, Activity) %>%
              summarize(across(everything(),mean))

```

The resulting dataframe was then exported as a csv file through write_csv and write.table.

```{r}
write_csv(merged_all, './Tidy_Data_Set.csv')
write.table(merged_all, file='./Tidy_Data_Set.txt', row.name=FALSE)
```

```{r, echo=FALSE}
merged_all
```